---
title: "LocalizationTaskAnalysis"
author: "Zeynep Bolluk, Rowan Mohamed, Esma Sakalli, Katharina Trant"
date: "2024-01-04"
output: html_document
---
# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Import all necessary packages.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(GGally)
library(ggcorrplot)
library(reshape2)
library(plotly)
```

# Data preprocessing

Set working directory, load all the localization data, clean it and combine it into one data frame. Also load the grasping data for later comparisons.

```{r Include = FALSE, eval=FALSE}
#this is for combining data files and taking only the last 3 blocks. If you're using the combined version, don't run this shell
# Set working directory to the folder containing the CSV files
setwd("C:/Users/Dell/Desktop/Courses/StudyProject/analysis/localizationTask/localization")
file_list <- list.files(pattern = "_localization*")
localization_data <- data.frame()

for (file in file_list) {
participant_id <- strsplit(file, "_")[[1]][1]
stimuli_file <- paste0(participant_id, "_stimuli.csv")
data_response <- read.csv(file, header = FALSE)
data_stimuli <- read.csv(stimuli_file, header = FALSE)

if (nrow(data_response) > 3) {
data_response <- tail(data_response, n = 3)
data_stimuli <- tail(data_stimuli, n = 3)
} else {
data_response <- tail(data_response, n = nrow(data_response))
data_stimuli <- tail(data_stimuli, n = nrow(data_stimuli))
}

response <- as.vector(t(data_response))
stimuli <- as.vector(t(data_stimuli))

block <- c()
for (i in 1:nrow(data_response)){
curr_block <- c(rep(i, ncol(data_response[i,])))
block <- c(block, curr_block)
}

dummy <- data.frame(response)
dummy$stimuli <- stimuli
dummy$participant_id <- participant_id
dummy$block <- block

localization_data <- rbind(localization_data, dummy)
}

localization_data$participant_id <- as.numeric(localization_data$participant_id)
localization_data <- localization_data[order(localization_data$participant_id), ]

# Add the code to select the last 3 blocks for each participant
localization_data <- localization_data %>%
group_by(participant_id) %>%
arrange(participant_id, block) %>%
slice_tail(n = 3) %>%
ungroup()

#exclude participant 23 in every data set as they have only auditory data
localization_data <- 
  filter(localization_data, 
         participant_id != 23)

#exclude participant 23 as they have only auditory, there was a testing with a 4th Block of tactile - exclude the 4th Block
combined_data <- 
  filter(combined_data, 
         participant_id != 23,
         block != 4)

# Save the combined data frame as a CSV file
row.names(localization_data) <- NULL
write.csv(localization_data, "localization_data.csv")

```

```{r include=FALSE}
# Set working directory to the folder containing the CSV files
setwd("/Users/katharinatrant/Desktop/Uni/Master/Semester 3/Study Project/Analysis/Data/Localization")

# load localization data
localization_data <- read.csv("localization_data.csv")

#load grasping data for comparison later
combined_data <- read.csv("combined.csv")

# Create inverted %in% function
`%ni%` <- Negate(`%in%`)

# fix the participant id
localization_data <- localization_data %>% group_by(participant_id) %>% mutate(participant_id = cur_group_id())

# Reorder columns putting 'participant_id' at the beginning
localization_data <- localization_data[, c("participant_id", setdiff(names(localization_data), "participant_id"))]

```

# Summary statistics & visualization

## Mean accuracies

```{r}
# Mean accuracy all trials
# Compute the number of lines where "response" and "stimuli" are equal
num_equal <- sum(localization_data$response == localization_data$stimuli)
# Compute the total number of lines
num_total <- nrow(localization_data)
# Compute the percentage of lines where "response" and "stimuli" are equal
percent_equal <- (num_equal / num_total) * 100
# Print the result
cat(sprintf("The accuracy over all trials is %.2f%%.", percent_equal))

# Mean accuracy all trials per participant
# Compute the percentage of lines where "response" and "stimuli" are equal for each participant
accuracy <- aggregate(localization_data$response == localization_data$stimuli, by = list(localization_data$participant_id), mean)
# Rename the columns of the result
colnames(accuracy) <- c("participant_id", "accuracy")
# Print the result
cat("\n\n")
summary(accuracy$accuracy)

# Boxplot mean accuracy
ggplot(accuracy, aes(y = accuracy)) +
  geom_boxplot() +
  labs(y = "Mean accuracy") +
  ggtitle("Boxplot of mean accuracies")
```
##Mean accuracy by block
```{r}
accuracybyblock <- aggregate(localization_data$response == localization_data$stimuli, by = list(localization_data$block),mean)

colnames(accuracybyblock) <- c('block_number', 'mean_accuracy') 

accuracybyblock$mean_accuracy_perc <- (accuracybyblock$mean_accuracy)* 100

#graph

accuracy_per_block <- accuracybyblock %>% ggplot(aes(x = block_number, y = mean_accuracy_perc)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Percentage of Mean Accuracy per block in Localization Task", 
    x = "\n Block number", 
    y = "Mean accuracy \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_x_continuous(breaks = c(1:8)) +
  scale_y_continuous(limits = c(0, 100))

accuracy_per_block


```

```{r}
# Turning values into percentage
accuracy$accuracy_perc <- accuracy$accuracy*100

# Create a new variable to mark participant with accuracy below 90
accuracy$color <- ifelse(accuracy$accuracy_perc < 90, "red", "default")

# Visualize the accuracies
ggplot(accuracy, aes(x = as.factor(participant_id), y = accuracy_perc, fill = color)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("red" = "#F8766D", "default" = "#545454")) +
  labs(title = "Mean accuracy by participant", x = "Participant ID", y = "Mean accuracy in percent") +
  coord_cartesian(ylim = c(40,100)) +
  geom_hline(yintercept = 90, color = "red", linetype = "dashed") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  guides(fill = FALSE)


```



## Direction confusion

Create a confusion matrix with counts and probabilities.

```{r}
# Create a table of counts for each combination of stimuli and response
count_table <- table(localization_data$stimuli, localization_data$response)

# Put "no response" column last
count_table <- count_table[,c(1,2,4,5,3)]

# Print the result
print(count_table)
```

¢
```{r}
# Convert the table into a data frame for ggplot2
ct_table <- count_table[, -5] # remove "no response"
count_df <- as.data.frame(melt(ct_table))

# Convert Var1 and Var2 to character
count_df$Var1 <- as.character(count_df$Var1)
count_df$Var2 <- as.character(count_df$Var2)

# Reorder the factor levels in Var2
count_df$Var2 <- factor(count_df$Var2, levels = c("down", "left", "right", "up", "no response"))

# Add a new column specifying whether or not each cell is on the diagonal
# Exclude "no response" column from the diagonal
count_df$on_diagonal <- count_df$Var1 == count_df$Var2 & count_df$Var2 != "no response"
count_df <- count_df %>% rename(count = value)

# Create the plot
confusion_counts <- ggplot(count_df, aes(Var2, Var1, fill = count)) +
  geom_tile() +
  scale_fill_gradient(low = "steelblue", high = "orange") +
  geom_text(aes(label = round(count, 3)), size = 4) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Response", y = "Direction command") +
  ggtitle("Count of vibration signal misinterpretation")
confusion_counts

#ggsave(paste0(SAVE,"confusion_counts.png"), plot = confusion_counts, dpi = 600)
```

```{r}
# Calculate how often 2 locations got confused.
up_right <- count_table["up","right"] + count_table["right","up"]
up_down <- count_table["up","down"] + count_table["down","up"]
up_left <- count_table["up","left"] + count_table["left","up"]
right_down <- count_table["right","down"] + count_table["down","right"]
right_left <- count_table["right","left"] + count_table["left","right"]
down_left <- count_table["down","left"] + count_table["left","down"]
noresponse <- sum(count_table[, "no response"])
right <- count_table["right", "right"]
left <- count_table["left", "left"]

cat(sprintf("Up and Right got confused %i times. \nUp and Down got confused %i times. \nUp and Left got confused %i times. \nRight and Down got confused %i times. \nRight and Left got confused %i times. \nDown and Left got confused %i times. \nIn %i trials the participant answered with no response. \nThe Down vibration motor was identified correctly most frequently, namely %i times. \nOverall participants correclty identified the Up motor the least. It was identified correctly %i times.", 
            up_right, up_down, up_left, right_down, right_left, down_left, noresponse, right, left))
```

Now for probabilities.

```{r}
# Create a confusion matrix by dividing the count table by the row sums
confusion_matrix <- prop.table(count_table, 1)

# Print the result
confusion_matrix
```

```{r}
# Convert the matrix into a data frame for ggplot2
confusion_df <- as.data.frame(melt(confusion_matrix))

# Convert Var1 and Var2 to character
confusion_df$Var1 <- as.character(confusion_df$Var1)
confusion_df$Var2 <- as.character(confusion_df$Var2)

# Reorder the factor levels in Var2
confusion_df$Var2 <- factor(confusion_df$Var2, levels = c("down", "left", "right", "up", "no response"))

# Add a new column specifying whether or not each cell is on the diagonal
# Exclude "no response" column from the diagonal
confusion_df$on_diagonal <- confusion_df$Var1 == confusion_df$Var2 & confusion_df$Var2 != "no response"

# Create the heatmap
confusion_matrix_perc <- ggplot(confusion_df, aes(Var2, Var1, fill = on_diagonal)) +
  geom_tile(aes(fill = on_diagonal)) +
  scale_fill_gradient(low = "steelblue", high = "orange") +
  geom_text(aes(label = round(value, 3)), size = 4) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Direction response", y = "Direction command") +
  ggtitle("Direction confusions in %") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
#confusion_matrix_perc

confusion_counts <- ggplot(count_df, aes(Var2, Var1, fill = count)) +
  geom_tile() +
  scale_fill_gradient(low = "steelblue", high = "orange") +
  geom_text(aes(label = round(count, 3)), size = 4) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Response", y = "Direction command") +
  ggtitle("Count of vibration signal misinterpretation")
confusion_counts

#ggsave(paste0(SAVE,"confusion_matrix_perc.png"), plot = confusion_matrix_perc, dpi = 600)


# create heatmap with ggcorrplot
confusion_mat <- confusion_matrix[, -5] # remove "no response"
confusions <- ggcorrplot(confusion_mat, lab=TRUE, digits=4, colors=c("#6D9EC1", "white", "#E46726"))
confusions

#ggsave(paste0(SAVE,"confusion_mat_cc.png"), plot = confusions, dpi = 600)
```

```{r}
# Calculate how often 2 locations got confused.
up_right <- confusion_matrix["up","right"] + confusion_matrix["right","up"]
up_down <- confusion_matrix["up","down"] + confusion_matrix["down","up"]
up_left <- confusion_matrix["up","left"] + confusion_matrix["left","up"]
right_down <- confusion_matrix["right","down"] + confusion_matrix["down","right"]
right_left <- confusion_matrix["right","left"] + confusion_matrix["left","right"]
down_left <- confusion_matrix["down","left"] + confusion_matrix["left","down"]
noresponse <- sum(confusion_matrix[, "no response"])
right <- confusion_matrix["right", "right"]
left <- confusion_matrix["left", "left"]

cat(sprintf("Up and Right got confused ~%.2f%% of the time. \nUp and Down got confused ~%.2f%% of the time. \nUp and Left got confused ~%.2f%% of the time. \nRight and Down got confused ~%.2f%% of the time. \nRight and Left got confused ~%.2f%% of the time. \nDown and Left got confused ~%.2f%% of the time. \nIn ~%.2f%% of trials the participant answered with no response. \nThe Down vibration motor was identified with the highest accuracy of ~%.2f%%. \nOverall participants identified the Up motor with the lowest acccuracy of ~%.2f%%.", up_right, up_down, up_left, right_down, right_left, down_left, noresponse, right, left))
```

## Predicting grasping performance with localization performance

### Reaction times

```{r}
# filter for only successes in tactile condition
mixed_data <- combined_data %>% filter(success=="success", condition=="tactile") %>% group_by(participant_id) %>% summarize(mean_time = mean(time)*1000)
mixed_data$accuracy <- accuracy$accuracy


# scatterplot: y = grasping_RTs, x = mean acc (first 3 blocks)
mixed_data %>%
  ggplot(aes(x = accuracy, y = mean_time, color= accuracy)) +
  geom_point() +
  labs(
    title = "Mean trial times per mean localization accuracy", 
    x = "\n Mean accuracy (%)", 
    y = "Mean trial times (ms) \n"
    ) +
  theme_linedraw() + 
  theme(plot.title = element_text(face = "bold"), legend.position = "right", legend.title = element_blank()) +
  scale_x_continuous(breaks = round(mixed_data$accuracy, 1))

```

```{r}
# barplot: y = mean RT, x = mean accuracy 
mixed_data %>% group_by(accuracy) %>% summarize(mean_time = mean(mean_time))
mixed_data %>% group_by(accuracy) %>% summarize(mean_time = mean(mean_time)) %>%
  ggplot(aes(x = accuracy, y = mean_time)) +
  geom_bar(stat = "identity", position = "dodge", colour="#CC9900", fill="#FFCC66") +
  labs(
    title = "Mean trial times per mean localization accuracy", 
    x = "Mean accuracy (%)", 
    y = "Mean trial times (ms) \n",
    ) +
  theme_linedraw() +
  scale_x_continuous(breaks = round(mixed_data$accuracy, 1)) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```
```{r}
# test for task performance correlation
cor.test(mixed_data$mean_time, mixed_data$accuracy)
```

